Confirmation of \srcfull{} as a long period radio transient

### 2025-05-05 paper/main.tex
Added another sentence or two to paper
Sync'ed paper with Overleaf version

\DeclareMathOperator{\erfcx}{erfcx}

@@

All observed pulses consist of a single burst, approximately Gaussian in shape, with the sole exception of the MeerKAT pulse observed on 2024-10-19, for which two bursts were observed separated by ${\sim}125\,$s.
Generally, the low-frequency (MWA) pulses tend
\begin{itemize}
        \item to be wider than their higher frequency counterparts,
              \item to exhibit an asymmetry similar to scattering, and
                    \item to have minimal pulse-to-pulse variation.
\end{itemize}

@@

To account for the asymmetry, we fit exponentially modified Gaussians to each pulse with the form,
\begin{equation}
  \begin{aligned}
        y(t) &= A \frac{\sigma}{\tau}\sqrt{\frac{\pi}{2}}
                   \exp\left(-\frac12 \left( \frac{t - \mu}{\tau} \right)^2 \right) \\
                            &\qquad \erfcx \left(\frac{1}{\sqrt{2}} \left(\frac{\sigma}{\tau} 
                             - \frac{t - \mu}{\tau} \right) \right),
  \end{aligned}
  \label{eqn:emg}
\end{equation}
where $A$, $\mu$, and $\sigma$ are the height, location, and scale of an unscattered Gaussian pulse, and the scattering time scale was fixed at
\begin{equation}
      \tau = 0.07\, \left(\frac{\nu}{\rm GHz}\right)^{-4} \, {\rm s},
\end{equation}
according to each observation's centre frequency.
A pulse stack containing all observed pulses and their fits is shown in Fig. \ref{fig:pulsestack}.

@@

Because the scattering time scale could not be reliably measured for all individual pulses, we obtained ToAs by assuming a fixed scattering time of $\tau_{\rm sc,1 GHz} = 0.07\,$s.
These ToAs were then barycentred and fit for both period and DM, resulting in the values given in Table \ref{tbl:ephemeris}.
We also tried fitting a timing model that included a spin-down parameter, and found that it was consistent with zero spin-down to within $1\sigma$.

\begin{table}
  \caption{Timing ephemeris for \src{}}
  \label{tbl:ephemeris}
  \begin{tabular}[lc]
    \hline
    Parameter & Value \\
    \hline
    Period (s) & $4186.3288 \pm 0.0003$ \\
    PEPOCH (MJD) & $59965.03783 \pm 0.00003$ \\
    DM (pc/cm$^{-3}$) & $771 \pm 27$ \\
    \hline
  \end{tabular}
\end{table}

@@

%\appendix

>>>>

\DeclareMathOperator{\erfcx}{erfcx}
\DeclareMathOperator{\erf}{erf}
\DeclareMathOperator{\emg}{emg}

\newcommand{\deriv}[2]{\frac{{\rm d}{#1}}{{\rm d}{#2}}}
\newcommand{\dd}[2]{\frac{{\rm d^2}{#1}}{{\rm d}{#2}^2}}
\newcommand{\ToA}[1]{{\rm ToA}_{\rm {#1}}}

@@

All observed pulses consist of a single burst, approximately Gaussian in shape, with the sole exception of the MeerKAT pulse observed on 2024-10-19, for which two bursts were observed separated by ${\sim}125\,$s.
Generally, the low-frequency (MWA) pulses tend
\begin{itemize}
    \item to be wider than their higher frequency counterparts,
    \item to exhibit an asymmetry resembling a scatter-broadened tail, and
    \item to have minimal pulse-to-pulse morphological variation.
\end{itemize}

@@

To account for the asymmetry, we fit exponentially modified Gaussians to each pulse with the form,
\begin{equation}
  \begin{aligned}
        \emg(t) &= A \frac{\sigma}{\tau}\sqrt{\frac{\pi}{2}}
                   \exp\left(-\frac12 \left( \frac{t - \mu}{\tau} \right)^2 \right) \\
                            &\qquad \erfcx \left(\frac{1}{\sqrt{2}} \left(\frac{\sigma}{\tau}
                             - \frac{t - \mu}{\tau} \right) \right),
  \end{aligned}
  \label{eqn:emg}
\end{equation}
where $A$, $\mu$, and $\sigma$ are the height, location, and scale of an unscattered Gaussian pulse, and the scattering time scale was fixed at
\begin{equation}
      \tau = 0.07\, \left(\frac{\nu}{\rm GHz}\right)^{-4} \, {\rm s},
\end{equation}
according to each observation's centre frequency.
A pulse stack containing all observed pulses and their fits is shown in Fig. \ref{fig:pulsestack}.

@@

Because the scattering time scale could not be reliably measured for all individual pulses, we obtained ToAs by assuming a fixed scattering time of $\tau_{\rm sc,1 GHz} = 0.07\,$s.
These ToAs were then barycentred and fit for both period and DM, resulting in the values given in Table \ref{tbl:ephemeris}.
The measured DM of $771 \pm 27\,{\rm pc}\,{\rm cm}^{-3}$ is slightly higher than, but still consistent with, the in-band measurement of \citetalias{2024MNRAS.535..909D}.

We also tried fitting a timing model that included a spin-down parameter, and found that it was consistent with zero spin-down to within $1\sigma$.

\begin{table}
  \caption{Timing ephemeris for \src{}}
  \label{tbl:ephemeris}
  \begin{tabular}{lc}
    \hline
    Parameter & Value \\
    \hline
    Period (s) & $4186.3288 \pm 0.0003$ \\
    PEPOCH (MJD) & $59965.03783 \pm 0.00003$ \\
    DM (pc/cm$^{-3}$) & $771 \pm 27$ \\
    \hline
  \end{tabular}
\end{table}

@@

\appendix

It can be shown that pulses scatter-broadened by a thin screen will appear to the observer as the original pulse convolved with a one-sided exponential kernel with an associated timescale $\tau$ [CITATION].
For intrinsically Gaussian pulses of scale $\sigma$, the result of this convolution is well described by the \textit{exponentially modified Gaussian} (EMG), as given in Eq. \eqref{eqn:emg}.

In the regime $\tau \ll \sigma$, the effect of scattering is negligible and the scattered pulse resembles the original pulse.
ToAs can be obtained in the usual way, i.e., using a matched filter ``template'' constructed from the average pulse profile.

On the other hand, in the regime $\tau \gg \sigma$, the leading edge of the EMG rises sharply (resembling the error function, $\erf$) and the trailing edge resembles the exponential kernel itself.
For this reason, pulse ToAs are typically mapped to the rising edge of highly scattered pulses, which closely approximates the centre of the original unscattered pulse.

In this appendix, we discuss the effect of scattering on measuring a DM from pulse observations if scattering is not taken into account.
As will be shown, the effect is most dramatic when $\tau \approx \sigma$, and in the case of wide, highly scattered pulses from LPTs like \src{}, can lead to a discrepant DM measurement of several hundreds of pm\,cm$^{-3}$.

We consider four methods for defining ToAs:
\begin{subequations}
(1) the position where the leading edge reaches half-maximum (LEHM),
\begin{equation}
    \emg(\ToA{LEHM}) = \frac{1}{2} A\exp\left(-\frac{1}{2}\left(\frac{\mu - t_m}{\sigma}\right)^2\right),
\end{equation}
where
\begin{equation*}
    t_m = \mu - \sqrt{2}\sigma\erfcx^{-1}\left(\frac{\tau}{\sigma}\sqrt{\frac{2}{\pi}}\right) + \frac{\sigma^2}{\tau^2}
\end{equation*}
is the mode of the EMG,
(2) the inflection point on the leading edge (IPLE),
\begin{equation}
    \left.\dd{\emg(t)}{t}\right|_{t = \ToA{IPLE}} = 0,
\end{equation}
(3) the position of the peak of the pulse (PEAK),
\begin{equation}
    \ToA{PEAK} = t_m,
\end{equation}
and (4) the maximum of the convolution of the pulse with a Gaussian profile with scale $\sigma$ (TMPL)
\begin{equation}
    \left.\deriv{}{t}\left(\emg(t) \ast \exp\left[ -\frac{1}{2} \left(\frac{t - \mu}{\sigma}\right)^2 \right]\right)\right|_{t = \ToA{TMPL}} = 0.
\end{equation}
The last definition is akin to template matching with a high-frequency (negligibly scattered) pulse profile.
\end{subequations}

None of the methods is entirely accurate in the presence of scattering.
We define the errors of the ToAs to be the difference between the measured ToA and the mean of the unscattered pulse,
\begin{equation}
    \Delta \ToA{} \equiv \ToA{} - \mu.
\end{equation}
For LEHM and IPLE, the ToA is measured to be earlier than $\mu$ ($\Delta{\rm ToA} < 0$); for PEAK and TMPL, later ($\Delta{\rm ToA} > 0$).
$\ToA{LEHM}$ and $\ToA{IPLE}$ will perform much better than $\ToA{PEAK}$ and $\ToA{TMPL}$ highly scattered pulses ($\tau \gg \sigma$), whereas the converse is true when scattering is minimal ($\tau \ll \sigma$).
All methods, however, perform relatively poorly in the intermediate regime when $\tau \approx \sigma$.

\begin{figure}[th]
    \centering
    \includegraphics[width=0.98\linewidth]{scattering_DM.png}
    \caption{Analysis of DM measurement errors if scattering is not properly taken into account. Panel (a) shows the errors measured for each defined type of ToA, normalised to the scale of the original pulse, $\sigma$. In panel (b), the predicted error in DM is shown if the ToA error (compared to a high-frequency ToA measurement) is erroneously attributed to dispersion. Panel (c) shows the rate of change of the ToA error with frequency, and panel (d) shows the error in DM measured if this rate of change (the ``slope'' of the pulse across the observed dynamic spectrum) is erroneously attributed to dispersion. In panels (c) and (d), the derived DM errors assume a pulse scale of $\sigma = 25\,$s and $\nu_s = 230\,$MHz, as estimated for \src{}. In all panels, the blue curves indicate \emph{negative} ToA errors (i.e. the measured ToAs are earlier than the pulse's unscattered mean, $\mu$), and the red curves indicate positive quantities.}
    \label{fig:scattering_DM}
\end{figure}

We computed numerically \citep[using SciPy's \texttt{root} function;][]{2020NatMe..17..261V} the errors for the four ToA types defined above, in the regime around $\tau \approx \sigma$.
These are shown in the panel (a) of Fig. \ref{fig:scattering_DM}, normalised to the scale of the pulse, $\sigma$.
The axis along the top shows the normalised timescale (inverted, with largest timescales on the left), while the axis along the bottom shows the equivalent frequencies normalised to $\nu_s$, defined as the frequency at which $\tau = \sigma$.
We have assumed a scattering index of $-4$, such that
\begin{equation}
    \frac{\tau}{\sigma} = \left(\frac{\nu}{\nu_s}\right)^{-4}.
\end{equation}

A set of ToAs may be converted to DM measurements in two ways: (1) by comparing the ToAs to those measured at much higher frequencies (e.g. with TMPL) where the scattering is known to be negligible, and (2) by measuring the slope of the pulse in the dynamic spectrum (for example, by measuring the ToAs in subbands across the observed frequency range).
Panel (c) of Fig. \ref{fig:scattering_DM} shows the expected slope measurements for the ToA errors given in the panel (a).

In the first case, the ToA error of a pulse measured at frequency $\nu$, if attributed erroneously to the effect of dispersion, will yield a discrepant DM of
\begin{equation}
    \Delta{\rm DM}_{\rm hi-freq} \approx \frac{\Delta\ToA{}\,\nu^2}{\mathcal{D}}.
\end{equation}
For the specific case of \src{}, for which we estimate $\sigma = 25\,$s and $\nu_s = 230\,$MHz, consistent with $\tau_{\rm sc,1 GHz} = 70\,$ms, the DM error can be as high as a few hundreds of pc\,cm$^{-3}$.
When using this DM method, it is better to use LEHM or IPLE at frequencies $\nu \lesssim \nu_s$, and PEAK or TMPL for $\nu \gtrsim \nu_s$; however, it should be remembered that LRHM and IPLE will \emph{underestimate} the DM, while PEAK and TMPL will \emph{overestimate} it.
All ToA definitions produce sizeable DM errors in the approximate range $\nu_s \lesssim \nu \lesssim 3\nu_s$.

The situation is not much improved for DMs determined via the slope of the pulse across the observed band, for which the scattering-induced slope is mapped to an (erroneous) equivalent DM measurement of
\begin{equation}
    \Delta{\rm DM}_{\rm in-band} \approx \deriv{{\rm ToA}}{\nu} \frac{\nu^3}{2\mathcal{D}}
\end{equation}
The calculated numbers for \src{} are illustrated in panel (d).
For this case, LEHM and IPLE are always preferred to PEAK and TMPL, but the magnitude of the derived DM error is still in the hundreds of pc\,cm$^{-3}$ at frequencies in the range $\nu_s \lesssim \nu \lesssim 3\nu_s$.

We confirm the order of magnitude of the above results by reporting a measured in-band DM for one of the MWA pulses at 185\,MHz, using the TMPL method to measure ToAs in 1.28\,MHz subbands from ${\sim}170$ to ${\sim}200\,$MHz, of ${\rm DM}_{\rm in-band} = 1221\pm257\,$pc\,cm$^{-3}$.
This exceeds the DM reported in this work by ${\sim}450\,$pc\,cm$^{-3}$, in excess of the prediction shown in panel (d), but consistent within measurement errors.
